PCA on Finnish Municipalities

### 2022-04-09 MSA-Project.Rmd
Added Scree Plot for PCA

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(summarytools)
library(pxweb)
library(ggplot2)
library(reshape2)

@@

@@

@@

ggplot(ages, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_bw() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Age Groups") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

@@

ggplot(living, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_bw() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Living Types") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

@@

```{r, echo=FALSE, fig.cap='Probability density estimates for Employment Groups', out.width='80%'}

ggplot(employment, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_bw() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Employment Groups") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

@@

```{r, echo=FALSE, fig.cap='Probability density estimates for Economic Sectors', out.width='80%'}

ggplot(living, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_bw() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Economic Sectors") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

@@

- Correlation heatmap here

@@

@@

>>>>

knitr::opts_chunk$set(echo = TRUE)
knitr::knit_hooks$set(optipng = knitr::hook_optipng)
library(tidyverse)
library(knitr)
library(kableExtra)
library(summarytools)
library(pxweb)
library(ggplot2)
library(reshape2)
library(ggcorrplot)

@@

<!-- \vspace{-1cm} -->

@@

<!-- \listoftables -->
<!-- \listoffigures -->

@@

ggplot(ages, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_minimal() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Age Groups") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

@@

ggplot(living, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_minimal() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Living Types") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

@@

```{r, echo=FALSE, fig.cap='Probability density estimates for employment groups', out.width='80%'}

ggplot(employment, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_minimal() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Employment Groups") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

@@

```{r, echo=FALSE, fig.cap='Probability density estimates for economic sectors', out.width='80%'}

ggplot(living, aes(x=value, fill=variable)) + geom_density(alpha = .25, adjust = .75) + theme_minimal() +
  scale_fill_manual( values = c("#dde318","#3dbc74","#440154")) +
  labs(x = "Percentage (%) of Population", y = "Density", fill = "Variable", title = "Economic Sectors") +
  theme(legend.position = "right", legend.text = element_text(size = 9))

@@

```{r, echo=FALSE, fig.dim=c(18,18), fig.cap="Heatmap of Pearson'n correlation between all variables", optipng = '-o7'}
corr <- cor(muni.latest[, -1])
ggcorrplot(corr, hc.order = TRUE, type = "full", lab = TRUE, title = "Heatmap of Correlations", lab_size = 3.5)

@@

## Scree Plot

```{r, echo=FALSE, fig.cap="Variance explained by each Princial Component. This type of a plot is also known as a Scree Plot", optipng = '-o7'}

muni.latest.pca <- muni.latest %>% column_to_rownames(var = "Area")
muni.pca <- prcomp(muni.latest.pca, scale. = T)

muni.pca.summary <- summary(muni.pca)
scree.df <- muni.pca.summary$importance[, 1:20] %>% t() %>% as.data.frame()
scree.df$PC <- 1:20
scree.df$Included <- c(rep("Yes",4), rep("No",16))
scree.df <- scree.df[, c(4,2,3,5)]
rownames(scree.df) <- NULL
colnames(scree.df) <- c("PC", "Prop.var", "Cum.prop", "Included")

ggplot(scree.df) +
  geom_bar(aes(x = PC, y = Prop.var, fill = Included), stat = "identity") +
  scale_fill_manual(values = c("#440154", "#3dbc74")) +
  geom_line(aes(x = PC, y = Cum.prop)) +
  geom_point(aes(x = PC, y = Cum.prop)) +
  geom_hline(yintercept = 0.75, linetype = 2) +
  annotate("text", x=10.5, y=0.78, label= "75% of variance explained", size = 3) +
  labs(x = "Principal Component", y = "Variance Explained (%)", title = "PCA Scree Plot") +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme_minimal()

## Biplot

@@

## Principal Components 3 and 4
